Terremotos Peninsula Iberica
========================== 

Terremotos producidos en la Peninsula Iberica 

******
#### Abril 2013
#### Guillermo Santos Garcia [@gsantosgo](http://twitter.com/gsantosgo)
#### This script is licensed under the GPLv2 license http://www.gnu.org/licenses/gpl.html
----------------------------------------------------------------
Note: This analysis was created on a Ubuntu 12.0.4 (RStudio 2.5.13). 
------

## Preliminaries

#### Set Working Directory
```{r working directory}
getwd()
WORKING_DIR <- "~/R/RStats/Data Analysis/"
DATASET_FILE <- "./data/terremotosSpain.rda"
FIGURES_DIR <- "./figures/"
setwd(WORKING_DIR)
getwd()
```


### Load libraries/data/create new variables

```{r loadData}

# Load libraries
#install.packages("ggmap")
library(ggmap)

# Load data
load(DATASET_FILE)

# Set Locale Time C 
lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")

# Define relevant variables - making numeric variables into factors when they should be.
terremotosSpainRaw$log10Depth <- log10(terremotosSpainRaw$Depth + 1)
terremotosSpainRaw$time <- strptime(terremotosSpainRaw$DateTime, "%Y-%m-%dT%H:%M:%S")

#Set default Locale Time 
Sys.setlocale("LC_TIME", lct)

## This is the data set we will use
terremotosSpain <- terremotosSpainRaw
```


------

## Exploratory analysis

### Get minimum and maximum times and date downloaded (Methods/Data Collection)

```{r, dependson="loadData"}
min(terremotosSpain$time)
max(terremotosSpain$time)
dateDownloaded
```

### Find number of missing values/check ranges (Results paragraph 1)

```{r, dependson="loadData"}
sum(is.na(terremotosSpain))
summary(terremotosSpain)
```

### Look at patterns over time 
```{r, dependson="loadData"}
plot(terremotosSpain$time,terremotosSpain$Magnitude,pch=19)
plot(terremotosSpain$time,terremotosSpain$Depth,pch=19)
```


### Look at distribution of magnitudes (Results paragraph 2)
```{r, dependson="loadData"}
mean(terremotosSpain$Magnitude < 3, na.rm=TRUE)
mean(terremotosSpain$Magnitude > 3 & terremotosSpain$Magnitude < 5, na.rm=TRUE)
```
Most earthquakes are small (< 3) or medium (>3 and < 5)


### Look at distribution of depths (Results paragraph 2)

```{r, dependson="loadData"}
hist(terremotosSpain$Depth,col="grey")
hist(terremotosSpain$log10Depth,col="grey")
```

### Map of Spain 
```{r, dependson="loadData"}
peninsulaSpain.map <- get_map( location = c(-9.5, 36, 3.5, 44), 
            color = "color", maptype = "roadmap")

# le añado puntos
ggmap(peninsulaSpain.map) +
  geom_point(aes(x = Longitude, y = Latitude,
                 size = Magnitude),
             data = terremotosSpain, colour = 'red',
             alpha = 0.2)
```