
Email: Spam or Not Spam  (CLASSIFICATION PROBLEM)
========================================================
This a classification problem (machine learning). 

More info:  
  http://archive.ics.uci.edu/ml/machine-learning-databases/spambase/  
  http://archive.ics.uci.edu/ml/machine-learning-databases/spambase/spambase.names  
  
  57 predictors  
  1 class (spam=1,email=0)   

******
#### May 9, 2013
#### Jose Maria Gomez Hidalgo [@jmgomez](http://twitter.com/jmgomez)
#### Guillermo Santos Garcia [@gsantosgo](http://twitter.com/gsantosgo)
#### This script is licensed under the GPLv2 license http://www.gnu.org/licenses/gpl.html
----------------------------------------------------------------

## Preliminaries
### Set Working Directory
```{r working directory}
getwd()
WORKING_DIR <- "C:/Users/gsantos/R/RStats/MadridJUG-DataMining"
setwd(WORKING_DIR)
getwd()
```

### Load libraries/data/create new variables

```{r loadData}
#install.packages(c("RColorBrewer","gridBase",ElemStatLearn","foreign","tree","rpart","maptree","class","ROCR)) 
library(RColorBrewer)
library(gridBase)
library(RColorBrewer)
library(ElemStatLearn)
library(foreign)
library(tree)
library(rpart)
library(maptree)
#install.packages("class") #not work in  2.15.2 
#library(class) #k Nearest Neighbors
library(e1071) # Support Vector Machine 
library(ROCR)

### Load data
DATASET <- spam 
```

### Set Color
```{r}
COLOR_SYSTEM <- "Set3"
COLORS <- brewer.pal(10,COLOR_SYSTEM)
PAL <- colorRampPalette(COLORS)
```

------

## Exploratory analysis

### Look at the data set

```{r, dependson="loadData"}
head(DATASET)
dim(DATASET)
nrow(DATASET)
ncol(DATASET)
colnames(DATASET)
sapply(DATASET[1,], class)
```


### Find number of missing values/check ranges

```{r, dependson="loadData"}
sum(is.na(DATASET))
summary(DATASET)
```


### Change Column Names 
```{r, dependson="loadData"}
newColNames <- c("word_freq_make",
"word_freq_address",
"word_freq_all",
"word_freq_3d",
"word_freq_our",
"word_freq_over",
"word_freq_remove",
"word_freq_internet",
"word_freq_order",
"word_freq_mail",
"word_freq_receive",
"word_freq_will",
"word_freq_people",
"word_freq_report",
"word_freq_addresses",
"word_freq_free",
"word_freq_business",
"word_freq_email",
"word_freq_you",
"word_freq_credit",
"word_freq_your",
"word_freq_font",
"word_freq_000",
"word_freq_money",
"word_freq_hp",
"word_freq_hpl", 
"word_freq_george",
"word_freq_650",
"word_freq_lab",
"word_freq_labs",
"word_freq_telnet",
"word_freq_857",
"word_freq_data",
"word_freq_415", 
"word_freq_85",
"word_freq_technology",
"word_freq_1999",
"word_freq_parts",
"word_freq_pm",
"word_freq_direct", 
"word_freq_cs",
"word_freq_meeting",
"word_freq_original",
"word_freq_project",
"word_freq_re",
"word_freq_edu",
"word_freq_table",
"word_freq_conference",
"char_freq_ch;",
"char_freq_ch(",
"char_freq_ch[",
"char_freq_ch!",
"char_freq_ch$",
"char_freq_ch#",
"capital_run_length_average",
"capital_run_length_longest",
"capital_run_length_total",
"spam")
length(newColNames)
colnames(DATASET) <- newColNames
colnames(DATASET)
```

### The Class Label
```{r, dependson="loadData"}
class(DATASET$spam)
levels(DATASET$spam)
```


### Numbers of Email in Dataset
```{r, dependson="loadData"}
result <- table(DATASET$spam)
numEmail <- result[["email"]]
numEmail
(numEmail/nrow(DATASET))*100
```

### Numbers of Spam in Dataset
```{r, dependson="loadData"}
numSpam <- result[["spam"]]
numSpam
(numSpam/nrow(DATASET))*100
```

### Numbers of Email vs. Numbers of Spam in DataSet
```{r, dependson="loadData"}
resTable <- table(DATASET$spam)
par(mfrow=c(1,2))
par(mar=c(5, 4, 4, 2) + 0.1) # increase y-axis margin.
plot <- plot(DATASET$spam,col=PAL(2), main="Email vs. Spam", ylim=c(0,4000), ylab="Examples Number")
text(x=plot,y=resTable+200, labels=resTable)
percentage <- round(resTable/sum(resTable)*100)
labels <- paste(row.names(resTable), percentage) # add percents to labels
labels <- paste(labels,"%",sep="") # ad % to labels
pie(resTable, labels=labels,col=PAL(2),main="Email vs. Spam")
```


### Average percentage of words or characters 
 Average percentage of words or characters in an email message
 equal to the indicated word or character. We have chosen the words and characters
 showing the largest difference between spam and email.
```{r, dependson="loadData"}
dataset.email <- sapply(DATASET[which(DATASET$spam == "email"), 1:54], function(x)ifelse(is.numeric(x), round(mean(x),2), NA))
dataset.spam <- sapply(DATASET[which(DATASET$spam == "spam"), 1:54], function(x)ifelse(is.numeric(x), round(mean(x),2), NA))

dataset.email.order <- dataset.email[order(-dataset.email)[1:10]]
dataset.spam.order <- dataset.spam[order(-dataset.spam)[1:10]]

par(mfrow=c(1,2))
par(mar=c(8, 4, 4, 2) + 0.1) # increase y-axis margin.
plot <- barplot(dataset.email.order, col=PAL(10), main='Email: Average Percentage', names.arg="", ylab="Percentage Relative (%)") 
#text(x=plot,y=dataset.email.order-0.1, labels=dataset.email.order, cex=0.6)
vps <- baseViewports()
pushViewport(vps$inner, vps$figure, vps$plot)
grid.text(names(dataset.email.order),x = unit(plot, "native"), y=unit(-1, "lines"),just="right", rot=50)
popViewport(3)

plot <- barplot(dataset.spam.order, col=PAL(10), main='Spam: Average Percentage', names.arg="", ylab="Percentage Relative (%)") 
#text(x=plot,y=dataset.spam.order-0.1, labels=dataset.spam.order, cex=0.6)
vps <- baseViewports()
pushViewport(vps$inner, vps$figure, vps$plot)
grid.text(names(dataset.spam.order),x = unit(plot, "native"), y=unit(-1, "lines"),just="right", rot=50)
popViewport(3)
```